---
layout: default
title: "Examples of workflows enabled by docker"
author: "Deepayan Sarkar"
editor_options: 
  chunk_output_type: console
---


```{r, include = FALSE}
knitr::opts_chunk$set(comment = "",
                      fig.path = "figures/dockereg-",
                      out.width = "100%",
                      fig.width = 12, fig.height = 8,
                      dev = "svglite", dev.args = list(pointsize = 12),
                      cache = TRUE,
                      warning = TRUE, error = FALSE,
                      cache.path = "~/knitr-cache/dockereg/")
options(width = 80)
```


## Introduction

One of the goals of the Epiconductor project is to make the analysis
of NHANES data simpler. To this end, it provides tools that can be
used to (a) download all public NHANES data and and load it into an
SQL database, and (b) make snapshots of this database available
locally through docker images. The primary goal in distributing the
pre-built docker images is to provide a ready-to-use tool for data
analysts who may not be conversant with database technologies. The
docker image thus includes R and RStudio, along with a version of the
`nhanesA` R package that is configured to obtain data from the SQL
database rather than the CDC website. We expect most users of the
docker image to use it through the `nhanesA` package. Such analyses
will run even outside the docker environment as well, albeit more
slowly as it will involve downloading data from the CDC website.

The potential uses of the docker image goes beyond this, however. As
NHANES collects and publishes data on a continuous basis, it is
natural to want to combine data across cycles. This is not
straightforward because a certain variable of interest may be present
in different tables in different cyles, or may not even be present in
all cycles. With all data locally available, it is possible to
efficiently check where variables are to be found (if at all), examine
differences over NHANES cycles, and detect a range of anomalies that
any analyst needs to be aware of, but would normally be difficult to
find. This vignette provides some illustrative examples, and describes
some relevant tools in the `phonto` package that could be useful to
practitioners.



- Combining data across cycles

- Consistency checks:

	- Across cycles: change in units

	- Within cycles: Variables appearing in multiple tables

- Variables that are potentially skipped


## Combining data across cycles


We start with a straightforward illustration of the process of
combining data across cycles, where we combine the demographic data
table from all cycles. NHANES table names typically have a suffix; the
suffixes `_A`, `_B`, `_C`,and so on generally correspond to NHANES
cycle years from 1999-2000, 2001-2002, 2003-2004, etc. However, it is
important to highlight that not every table strictly adheres to this
naming convention. For instance, while `DEMO_B` and `DEMO_C` are
associated with the 2001-2002 and 2003-2004 cycles, respectively, the
corresponding table for the 1999-2000 cycle is named 'DEMO', without
the `_A` suffix. While this pattern holds for most tables, certain
tables such as `SSAFB_A` and `SSANA_A` from the 1999-2000 cycle do
include the `_A` suffix. To assist users in navigating these
variations, the `nhanesA` package includes the
`nhanesSearchTableNames()` function, which allows users to easily
locate all table names containing a specific string, thus simplifying
the process of identifying relevant table names.

```{r alldemo}
library(nhanesA)
demo_all <- nhanesSearchTableNames("DEMO")
demo_all
```

The last table in this list merits special mention. During the
2019-2020 cycle, data collection was disrupted by the COVID-19
pandemic. Therefore, the partial 2019-2020 data (herein 2019-March
2020 data) were combined with data from the previous cycle (2017-2018)
to create a nationally representative sample covering 2017-March 2020.
These data files have the same basic file name, e.g., `DEMO`, but add
the prefix `P_`. These 'pre-pandemic' files require special handling
and the CDC has provided substantial guidance as well as updated
survey weights.

We can now collect all these datasets using `nhanes()`
function. Normally, this process is likely to be somewhat slow as
several files will need to be downloaded from the CDC website. Within
the docker environment, however, it should be almost instantaneous.

```{r getalldemo, error = TRUE}
all_demo_data <- sapply(demo_all, nhanes, simplify = FALSE)
object.size(all_demo_data) # ~45 MB
sapply(all_demo_data, dim)
```

The first row in the output above gives the number of participants in
each cycle, and the second row denotes the number of variables in the
corresponding `DEMO` table. We can see that each cycle has around
10,000 participants, who are unique across cycles. Note, however, that
the larger number of participants in the `P_DEMO` dataset is
misleading, because many of these participants are actually from the
previous cycle as described above. We will drop this table before
combining the remaining datasets.

The differing number of variables across cycles indicate that
variables are not measured consistently across cycles. In fact, many
variables included in the `DEMO` table in the first cycle were
subsequently included in other tables, and others have been dropped
altogether or added. We can make a list of the variables that are
common to all `DEMO` tables, and combine the corresponding data
subsets together, as follows.

```{r combinedemo}
all_demo_data <- head(all_demo_data, -1)
common_vars <- lapply(all_demo_data, names) |> Reduce(f = intersect)
common_vars
demo_combined <-
    lapply(all_demo_data, `[`, common_vars) |>
    do.call(what = rbind) |>
    transform(cycle = substring(SDDSRVYR, 8, 17))
dim(demo_combined)
```

The combined dataset can be analysed further using standard tools. For
example, the following code uses the `lattice` package to summarize
the number of participants by recorded ethnicity and gender by cycle.

```{r demoplot,fig.width=12, fig.height=7, fig.caption=""}
library("lattice")
demo_combined |>
    xtabs(~ cycle + RIAGENDR + RIDRETH1, data = _) |>
    array2DF() |>
    dotplot(Value ~ cycle | RIAGENDR,
            groups = RIDRETH1,
            layout = c(1, 2), type = "b",
            par.settings = simpleTheme(pch = 16),
            auto.key = list(columns = 3))
```


One must be cautious when combining data across
cycles, because the NHANES data are sometimes inconsistent in
unexpected ways. As a simple example, consider the `DMDEDUC3`
variable, which records education level of children and youth. The
following code illustrates that the values of this variable have
inconsistent capitalization in different cycles.

```{r changes}
xtabs(~ cycle + DMDEDUC3, demo_combined)[, 1:4]
```


## Cross-cycle Consistency checks using variable codebooks

In our experience, inconsistencies such as the change in
capitalization described above occur quite often, and in a variety of
different ways, requiring attention to detail when combining data from
across cycles. These inconsistencies are not necessarily mistakes, as
NHANES questionnaires and variables are often modified from cycle to
cycle. The primary source that must be consulted to identify such
inconsistencies are the per-table documentation, and in particular the
variable codebooks giving details of how each variable is recorded.

The NHANES database contains the variable codebooks for all tables
across all cycles in a single database table called
`Metadata.VariableCodebook`. Once imported into R, this information
can be manipulated in various ways to glean information of
interest. See the [codebook
diagnostics](https://ccb-hms.github.io/phonto/vignettes/diagnostics-codebook.html)
vignette for a different set of diagnostics that are enabled by having
easy access to this information.

```{r}
library("phonto")
dim(all_cb <- nhanesQuery("select * from Metadata.VariableCodebook"))
## dim(all_cb <- metaData("Codebook")) # alt interface
all_cb <- dplyr::filter(all_cb, !startsWith(TableName, "P_")) # skip pre-pandemic tables
```

An analyst would typically be interested in some specific variables
that are relevant to their study. The first step is to identify how
many cycles these variables were recorded in.  To this end, we may
start by examining the number of _tables_ each variable appears in
across all cycles of continuous NHANES.

```{r}
var_freq <- 
    all_cb[c("Variable", "TableName")] |> unique() |>
        xtabs(~ Variable, data = _) |>
        sort(decreasing = TRUE)
table(var_freq)
```

Not surprisingly, many variables appear only once or twice across all
cycles, and for such variables combining data across cycles would not
be interesting. Variables that appear in multiple cycles, but only
once per cycle, may usually be merged as in the `DEMO` example above,
but note that they may appear in different tables. One must be careful
about ensuring that the variable is measuring the same quantity in all
cycles. This is usually true, but not always. For example, the range
of values for the `LBCBHC` variable in different cycles seem to
exhibit some discrepancy.

```{r}
subset(all_cb, Variable == "LBCBHC")[1:5]
```

More useful information about the variables may be obtained from the
`Metadata.QuestionnaireVariables` table in the database, which
contains one row for each variable in each table containing its
description, target group, etc., obtained from the HTML documentation
of NHANES tables.

```{r}
dim(all_var <- phonto::nhanesQuery("select * from Metadata.QuestionnaireVariables"))
## dim(all_var <- metaData("Variables"))
subset(all_var, Variable == "LBCBHC")[1:5]
```

This shows that the unit of measurement was changed from the
2009--2010 cycle, explaining the discrepancy. Without a careful check,
such changes may be overlooked, leading to errors in interpretation.

It is not easy to systematically detect such changes without manual
inspection of variables of interest. One way to shortlist possible
candidate variables are to identify those for whom the `Description`
or `SasLabel` field has changed. Unfortunately, such changes happen
frequently for completely benign reasons, leading to many false
positives.


## Within-cycle consistency

Somewhat more surprisingly, several variables appear
in more tables than there are cycles, which means that they must
appear in multiple tables within the same cycle. The following
variables appear in more than 20 tables.

```{r}
var_freq[ var_freq > 20 ]
```

For such variables, selecting the corresponding subset of `all_cb`
shows all entries in the codebook tables for that variable, across all
tables and cycles. For example,

```{r}
subset(all_cb, Variable == "PHAFSTMN")[1:5]
```


Inspection of this table shows that the `PHAFSTMN` variable was
initially recorded in the `PH` table for the first three cycles, after
which it was recorded in three different tables (`FASTQX`, `GLU`, and
`OGTT`) for several cycles, before being dropped again from the latter
two tables. It is natural to wonder whether all these tables contain
the same data. This can only be verified by comparing the actual data,
which we will not do for this example, but some hints are provided by
the data counts included in the codebook. For example, for the
2005--2006 cycle, we have

```{r}
subset(all_cb, Variable == "PHAFSTMN" & endsWith(TableName, "_D"))[1:5]
```

From the variable metadata table, we see that 

```{r}
subset(all_var, Variable == "PHAFSTMN" & endsWith(TableName, "_D"))[c(1, 2, 4, 5)]
```

While not definitive, this suggests that the `PHAFSTMN` variable
measures the same quantity in all three tables, and the difference in
number of observations may be due to the difference in target age
group.

Even if a preliminary inspection suggests no obvious problems, one
should verify by comparing the actual recorded data. For example,
consider the `ENQ100` variable, which appears in both `ENX` and `SPX`
tables.


```{r}
subset(all_var, Variable == "ENQ100" & endsWith(TableName, "_E"))[c(1, 2, 4, 5)]
merge(nhanes("ENX_E")[c("SEQN", "ENQ100")],
      nhanes("SPX_E")[c("SEQN", "ENQ100")], by = "SEQN") |>
    xtabs(~ ENQ100.x + ENQ100.y, data =_, addNA = TRUE)
```

Comparing the records in the two tables after matching by `SEQN`, the
participant identifier, we see that even though most records are
consistent, several records with `Yes` or `No` answers in the `SPX`
tables are recorded as `NA` (missing) in the `ENX` tables.


A more egregious example, where the same variable is clearly measuring
two different things, is provided by the `LBXHCT` variable.

```{r}
subset(all_cb, Variable == "LBXHCT" & endsWith(TableName, "_H"))[1:5]
subset(all_var, Variable == "LBXHCT" & endsWith(TableName, "_H"))[c(1, 2, 3, 4, 5)]
merge(nhanes("CBC_H")[c("SEQN", "LBXHCT")],
      nhanes("COT_H")[c("SEQN", "LBXHCT")], by = "SEQN") |> head()
```


## Quality control reports

Getting a comprehensive picture of the issues relevant for a
particular variable may take some work, as seen in the above
examples. However, much of the preliminary work can be automated, and
the `qc_var()` function in the `phonto` package dowa precisely this.

The report for the `LBCBHC` variable can be generated using the
following call.

```{r}
qc_var("LBCBHC")
```

Similarly, the reports for `PHAFSTMN` (where the issues flagged are
relatively benign) and `LBXHCT` (where they are not) are obtained as
follows.

```{r}
qc_var("PHAFSTMN")
qc_var("LBXHCT")
```

As we can see, these reports give an incomplete but suggestive initial
list of potential issues, which can be followed up on by the analyst
if necessary.

Other useful checks that have not yet been implemented but are planned
include:

* Special codes in numeric variables: For numeric variables, special
  codes are sometimes used to indicate unusual values. A common use is
  to have special codes to disambiguate various types of missing
  responses (e.g., `"Refused"` and `"Don't know"`). Some of these are
  already converted to `NA` values by default, but others may remain.

* One particular instance of special codes represent data coarsening,
  that is, clubbing of numeric values over a range (e.g., ages over 80
  are usually replaced by 80).

The plan is to report the presence of such codes in variable QC
reports.

## Variables that are potentially skipped

Some variables in a table may be skipped depending on the response to
a previous question. This information is recorded in the variable
codebooks. For example, in table `WHQ`, if the response to `WHD060` is
"Yes", then subsequent questions are skipped until `WHQ090` is
reached.

```{r}
nhanesCodebook("WHQ", "WHD060")
```

This information, while useful, does not directly tell us _which_
variables were skipped. This may be important for the analyst to know,
because a variable that was skipped is recorded as a missing value,
whereas we may actually have additional information based on whether
it was skipped or not. In the above example, the immediately following
variable is `WHQ070`.

```{r}
nhanesCodebook("WHQ", "WHQ070")
```

This codebook gives no indication of whether this question could have
been skipped for some participants. However, from the previous
codebook, was can infer that out of the 518 missing values, 508 were
presumably from those that said "Yes" to `WHD060`, and these are
qualitatively different from the remaining 10. While the question of
_how_ to address such situations is for the analyst to answer, it
would be useful to flag variables that are potentially skipped.

The `get_skip_info()` function in the `phonto` package reports the
variables that are potentially skipped in any given table. The
following code applies it on the `WHQ` table.

```{r}
get_skip_info("WHQ")
```


## Searching variables

Having access to all variable descriptions in a single dataset makes
search operations convenient. For example, 

```{r}
all_var |> within({ Description <- tolower(Description) }) |>
    subset(endsWith(TableName, "_C") &
           (grepl("hypertension", Description) |
            grepl("blood pressure", Description)),
           select = c(1, 2, 5))
```

Information in this table can be supplemented using the
`nhanesTableSummary()` function, which computes further information
such as the variable type and number of non-missing observations; see
[this vignette](search-tables.html) for more details and links to
publicly accessible online search interfaces.


