---
layout: default
title: "Introduction to Continuous NHANES"
author: "Deepayan Sarkar"
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
library(knitr)
opts_chunk$set(collapse = FALSE, autodep = TRUE, 
               comment = "", warning = TRUE, message = TRUE, prompt = FALSE,
               fig.path = "figures/nhanes-intro-",
               out.width = "100%",
               fig.width = 12, fig.height = 8,
               dev = "svglite", dev.args = list(pointsize = 12),
               cache = TRUE,
               cache.path = "~/knitr-cache/nhanes-intro/")
options(warnPartialMatchDollar = FALSE, width = 90)
library(lattice)
library(kableExtra)
```

The [National Health and Nutrition Examination
Survey](https://www.cdc.gov/nchs/nhanes/about_nhanes.htm) (NHANES) is
a program of the National Center for Health Statistics (NCHS), which
is part of the US Centers for Disease Control and Prevention (CDC). It
measures the health and nutritional status of adults and children in
the United States in a series of surveys that combine interviews and
physical examinations.

Although the program began in the early 1960s, its structure was
changed in the 1990s.  Since 1999, the program has been conducted on
an ongoing basis, where a nationally representative sample 
of about 5,000 persons (across 15 counties) is examined each year,
with public-use data released in two-year cycles. This phase of the
program is referred to as [continuous
NHANES](https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/).

The NHANES interview includes demographic, socioeconomic, dietary, and
health-related questions. The examination component consists of
medical, dental, and physiological measurements, as well as laboratory
tests administered by highly trained medical personnel. Although the
details of the responses recorded vary from cycle to cycle, there is a
substantial amount of consistency, making it possible to compare data
across cycles.  Sampling weights are provided along with demographic
details for each participant; see the NHANES [analytic
guidelines](https://wwwn.cdc.gov/nchs/nhanes/analyticguidelines.aspx)
for details.


# Public-use data: web resources

NHANES makes a large volume of data available for download. However,
rather than a single download, these data are made available as a
number of separate SAS transport files, referred to as "data files",
for each cycle. Each such data file contains records for several
related variables. A comprehensive list of data files available for
download is available
[here](https://wwwn.cdc.gov/Nchs/Nhanes/search/DataPage.aspx), along with
subsets broken up into the following "components": 
[Demographics](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Demographics),
[Dietary](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Dietary),
[Examination](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Examination),
[Laboratory](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Laboratory), and
[Questionnaire](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Questionnaire).

For each data file listed in these tables, a link to a Doc File (which
is an HTML webpage describing the data file) and a link to a SAS
transport file is provided. An additional list of limited access data
files are documented
[here](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=LimitedAccess),
but the corresponding data file download links are not available.

A table of _variables_ is separately available for each component, and
gives more detailed information about both the variables and the data
files they are recorded in, although these tables do not provide
download links directly: 
[Demographics](https://wwwn.cdc.gov/nchs/nhanes/search/variablelist.aspx?Component=Demographics),
[Dietary](https://wwwn.cdc.gov/nchs/nhanes/search/variablelist.aspx?Component=Dietary),
[Examination](https://wwwn.cdc.gov/nchs/nhanes/search/variablelist.aspx?Component=Examination),
[Laboratory](https://wwwn.cdc.gov/nchs/nhanes/search/variablelist.aspx?Component=Laboratory),
[Questionnaire](https://wwwn.cdc.gov/nchs/nhanes/search/variablelist.aspx?Component=Questionnaire).

In addition, a [search interface](https://wwwn.cdc.gov/nchs/nhanes/search/) is also available.

For reasons [not
specified](https://wwwn.cdc.gov/nchs/nhanes/sasviewer.aspx), NHANES
releases data files as SAS transport files, and provides links to
proprietary Windows-only software that can supposedly be used to
convert these files to CSV files.


# Public-use data: R resources

One of the goals of the Epiconductor project is to provide and
document an alternative access path to NHANES data and documentation
_via_ the R ecosystem. It builds on the
[__nhanesA__](https://cran.r-project.org/package=nhanesA) R package,
along with utilities such as SQL databases and docker, to enable
efficient and reproducible analyses of NHANES data.

## The __nhanesA__ package

The [__nhanesA__](https://github.com/cjendres1/nhanes) package provides
a user-friendly interface to download and process data and
documentation files from the NHANES website. To use the utilities in
this package, we first need to know a few more details about how
NHANES data and documentation are structured.

Each available data file, which we henceforth call an NHANES _table_,
can be identified uniquely by a name. Generally speaking, each
public-use table has a corresponding data file (a SAS transport file,
with extension `xpt`) and a corresponding documentation file (a
webpage, with extension `htm`). The URLs from which these files can be
downloaded can usually be predicted from the table name, and the
_cycle_ it belongs to. Cycles are typically of 2-year duration,
starting from `1999-2000`.

Although there are exceptions, a table that is available for one cycle
will typically be available for other cycles as well, with a suffix
appended to the name of the table indicating the cycle. To make these
details concrete, let us use the `nhanesManifest()` function in the
__nhanesA__ package to download the [list of available
tables](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx) and
look at the names and URLs for the `DEMO` data files, which contain
demographic information and sampling weights for each study
participant.


```{r}
library(nhanesA)
manifest <- nhanesManifest("public")
manifest <- manifest[order(manifest$Table), ]
subset(manifest, startsWith(Table, "DEMO"))
```

The __nhanesA__ package allows both data and documentation files to be
accessed, either by specifying their URL explicitly, or simply using
the table name, in which case the relevant URL is constructed from
it. For example,

```{r}
demo_b <- nhanesFromURL("/Nchs/Nhanes/2001-2002/DEMO_B.XPT", translated = FALSE)
demo_c <- nhanes("DEMO_C", translated = FALSE)
```

```{r}
str(demo_b[1:10])
str(demo_c[1:10])
```

The data in these files appear as numeric codes, and must be
interpreted using codebooks available in the documentation files,
which can be parsed as follows.

```{r}
demo_b_codebook <-
    nhanesCodebookFromURL("/Nchs/Nhanes/2001-2002/DEMO_B.htm")
demo_b_codebook$RIDSTATR 
demo_b_codebook$RIAGENDR
```

By default, the data access step converts the raw data into more
meaningful values using the corresponding codebook.

```{r}
demo_c <- nhanes("DEMO_C", translated = TRUE)
str(demo_c[1:10])
```

Further analysis can be performed on these resulting datasets which
are regular R data frames.

Other tools that make it easier to work with these datasets by
creating a local database, or creating a search interface, are
described elsewhere. We conclude this document with a brief look at
how frequently NHANES data files are published and / or updated, based
on the information contained in the table manifest.


## Frequency of NHANES data releases

Recall from above that the NHANES table manifest includes a
`Date.Published` column.  This allows us to tabulate NHANES data
release dates. We expect that bulk releases of tables happen all
together, generally in two year intervals, while some tables may be
released or updated on a as-needed basis.

The release information (available by month of release) can be
summarized by tabulating the `Date.Published` field:
```{r}
xtabs(~ Date.Published, manifest) |> sort() |> tail(20)
```

Parsing these dates systematically, we get

```{r}
pubdate <- manifest$Date.Published
updates <- startsWith(pubdate, "Updated")
datesplit <- strsplit(pubdate, split = "[[:space:]]")
datesplit[updates] <- lapply(datesplit[updates], "[", -1)
pub_summary <-
    data.frame(updated = updates,
               year = sapply(datesplit, "[[", 2) |> as.numeric(),
               month = sapply(datesplit, "[[", 1) |> factor(levels = month.name))
```

Although there are a few too many months, we can plot the number of
releases + updates by month as follows.

```{r, bymonth, fig.cap = ""}
pubfreq <- xtabs(~ interaction(month, year, sep = "-") + updated, pub_summary)
npub <- rowSums(pubfreq)
npub.date <- as.Date(paste0("01", "-", names(npub)), format = "%d-%B-%Y")
xyplot(npub ~ npub.date, type = "h", grid = TRUE,
       xlab = "Month", ylab = "Number of tables published / updated") +
    latticeExtra::layer(panel.text(x[y > 30], y[y > 30],
                                   format(x[y > 30], "%Y-%m"),
                                   pos = 3, cex = 0.75))
```


We can also plot the release / update frequency by year as follows.

```{r, byyear, fig.cap = ""}
xtabs(~ year + updated, pub_summary) |>
    barchart(horizontal = FALSE, ylab = "Number of tables",
             auto.key = list(text = c("Original", "Update"), columns = 2),
             scales = list(x = list(rot = 45)))
```

A full table of number of releases by month is given by the following,
showing that there is at least one update almost every month.

```{r}
pubfreq0 <- pubfreq[rowSums(pubfreq) > 0, , drop = FALSE]
pubfreq0
```


