---
layout: default
title: "Searching NHANES Variables and Tables"
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
library(knitr)
opts_chunk$set(collapse = FALSE, autodep = TRUE, 
               comment = "", warning = TRUE, message = TRUE, prompt = FALSE,
               fig.path = "figures/search-tables-",
               out.width = "100%",
               fig.width = 12, fig.height = 8,
               dev = "svglite", dev.args = list(pointsize = 12),
               cache = TRUE,
               cache.path = "./knitr-cache/search-tables/")
options(warnPartialMatchDollar = FALSE, width = 200)
```

A broad overview of data collected as part of continuous NHANES is
described in the [survey content
brochure](https://wwwn.cdc.gov/nchs/data/nhanes/survey_contents.pdf),
and the NHANES website has a [variable search
interface](https://wwwn.cdc.gov/nchs/nhanes/search/default.aspx) as
well. This document describes how local interactive searches can be
performed using downloaded and processed variable metadata
information.


# Manifest of available variables

The NHANES data consist of multiple tables over multiple cycles, each
with a set of recorded variables. The variable names themselves are
cryptic, and not useful as search terms. Fortunately, the NHANES
website provides comprehensive details about these variables in the
form of tables, grouped by components such as
[Demographics](https://wwwn.cdc.gov/nchs/nhanes/search/variablelist.aspx?Component=Demographics),
[Laboratory](https://wwwn.cdc.gov/nchs/nhanes/search/variablelist.aspx?Component=Laboratory),
etc.

The __nhanesA__ package can download these tables and make the information
contained in them available as a data frame using the
`nhanesManifest()` function.

```{r}
library(nhanesA)
library(phonto)
varmf <- nhanesManifest("variables")
```

This downloads and combines several large web pages, so it may take a
little time to run. Add `verbose = TRUE` to see some indication of
what is happening.

While useful, this table has limited information. The most useful
component, the variable description, sometimes (but not always)
records instructions and can be quite long. 


```{r}
sum(nchar(varmf$VarDesc) == 0)
```

After reordering the remaining rows by the length of the descriptions,
the first few rows are given by

```{r}
varmf <- subset(varmf, nzchar(varmf$VarDesc))
varmf <- varmf[ order(nchar(varmf$VarDesc)), ]
head(varmf)
```

The last few rows have considerably longer descriptions.

```{r}
unique(tail(varmf, 30)$VarDesc)
```

More useful and concise descriptions are available in the per-table
documentation, which can be accessed using the `nhanesCodebook()`
function and summarized using the `nhanesTableSummary()` function. For
example, a summary of the `DEMO_C` table if given by the following.

```{r}
nhanesTableSummary("DEMO_C", use = "both")
```

This collects information from the codebook and supplements it with
information from the actual data, such as the number of non-missing
observations. Combining such information across all available tables
is generally time-consuming, but is considerably easier when the data
are available in the form of a database. 

# Summarizing variables using the SQL database

<!-- 

These are useful: should we have a R function to get these in phonto?

phonto::nhanesQuery("select * from Metadata.QuestionnaireDescriptions") |> str()
phonto::nhanesQuery("select * from Metadata.QuestionnaireVariables") |> str()
phonto::nhanesQuery("select * from Metadata.VariableCodebook") |> str()

-->

Basic information about all tables can be obtained using
`nhanesManifest("public")`, but also from the database using a SQL
query as follows.

```{r}
tableDesc <- nhanesQuery("select * from Metadata.QuestionnaireDescriptions")
subset(tableDesc, BeginYear == 1999, select = -c(DocFile, DataFile)) |> head(10)
```

Having a searchable list of these table descriptions can by itself be
useful. Such a table is available
[here](../tables/table-summary.html), and is created using [this R
script](../scripts/generate-table-summary.R). Instead of having one
row per table, this list combines multiple cycles for the same _base_
table name and description, and includes the dimensions of each
dataset.

We can now apply the `nhanesTableSummary()` function on each of these
tables. The essential code for this, accounting for the possibility of
errors, is

```{r, eval=FALSE}
table_summary <- function(x) {
    try(nhanesTableSummary(x, use = "both"), silent = TRUE)
}
table_details <- sapply(sort(tablemf$TableName), table_summary, simplify = FALSE)
nhanesVarSummary <-
    table_details[!sapply(table_details, inherits, "try-error")] |>
        do.call(what = rbind)
```

```{r, echo=FALSE}
load("../variable_summary.rda")
```

This results in a fairly large data frame.

```{r}
str(nhanesVarSummary)
```

However, as variables are repeated for each cycle, the number of
_unique_ variables is considerably smaller, especially if we split
them according to the component or data group.

```{r}
rownames(tableDesc) <- tableDesc$TableName
nhanesVarSummary$DataGroup <- tableDesc[nhanesVarSummary$table, "DataGroup"]
unique(nhanesVarSummary[c("varname", "DataGroup")]) |>
    xtabs(~ DataGroup, data = _)
```

Having variable-level information in a data frame like this allows us
to perform interactive searches in R quite easily. For example, we
could search for variables related to blood pressure or hypertension recorded in the 
third cycle as follows.

```{r}
nhanesVarSummary <- within(nhanesVarSummary, label <- tolower(label))
subset(nhanesVarSummary, endsWith(table, "_C") &
                         (grepl("hypertension", label) |
                          grepl("blood pressure", label)))
```

As with the table descriptions, information about these variables can
also be summarized and presented as searchable HTML tables. Sample
code for doing this is available
[here](../scripts/generate-variable-summary.R), producing the
following tables, grouped by data group to keep the table sizes
moderate.

- [Demographics](../tables/variable-summary-demographics.html)

- [Questionnaire](../tables/variable-summary-questionnaire.html)

- [Examination](../tables/variable-summary-examination.html)

- [Laboratory](../tables/variable-summary-laboratory.html)

- [Dietary](../tables/variable-summary-dietary.html)

