---
title: "quck_start"
output: html_document
# output: rmarkdown::html_vignette
# vignette: >
  # %\VignetteIndexEntry{quck_start}
  # %\VignetteEngine{knitr::rmarkdown}
  # %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(phonto)
library(DT)
```

### searchTableByName

We can do the following can find the tables with string "BPX".
```{r}
res = searchTableByName("BPX")
unique(res$TableName)
datatable(res)
```

You may notice that the table name of the last row is different from the rest of the row, and you may want to try something like the one below to get rid of it. However, the following method would need to include the first row in the above result. Therefore, we need to be careful of the patterns we pass in the function; otherwise, we may miss some of the data because the name convention of NHANES does not always keep consistent.  
```{r}
res = searchTableByName("BPX[_]")
datatable(res)
```
### unionQuery and PHESANT-like
`r_unique` and `r_zero` denote unique ration compute as number of unique value and zero dived by number of total records.

```{r}
cols = c("DR1DAY","DRQSPREP","DR1TNUMF","DR1TKCAL","DR1TPROT","DR1TCARB","DR1TSUGR","DR1TFIBE","DR1TTFAT","DR1TSFAT","DR1TMFAT","DR1TPFAT","DR1TCHOL","DR1TATOC","DR1TRET","DR1TVARA","DR1TBCAR","DR1TCRYP","DR1TLZ","DR1TVB1","DR1TVB2","DR1TNIAC","DR1TVB6","DR1TFOLA","DR1TFA","DR1TFF","DR1TFDFE","DR1TVB12","DR1TVC","DR1TVK","DR1TCALC","DR1TPHOS","DR1TMAGN","DR1TIRON","DR1TZINC","DR1TCOPP","DR1TSODI","DR1TPOTA","DR1TSELE","DR1TMOIS","DR1TS040","DR1TS080","DR1TS120","DR1TS160","DR1TS180","DR1TM161","DR1TM181","DR1TM201","DR1TP183","DR1TP204")
diet = unionQuery("DietaryInterviewTotalNutrientIntakesFirstDay",cols)
phs_dat = phesant(diet)
datatable(phs_dat$phs_res)
```
