---
title: "Query Data"
date: "Updated on : `r date()`"
output: html_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Setup functions


```{r setup,warning=FALSE,message=FALSE}
library(phonto)
```


### Check the data availability over the years
In the following matrix, 0 means the tables have no data in that year, and 1 means they have data in corresponding years.

```{r years,warning=FALSE,message=FALSE}
mtx <- check_data()
knitr::kable(mtx)
```



### Query and joint data
Based on the above, it seems we can joint PhytoestrogensUrine, PhthalatesUrine, and MetalsUrine. However, it will get an empty data frame if we do that.


```{r query1,warning=FALSE,message=FALSE}
df <- query_joint_data(config = "./phenotypes2.json")
nrow(df)
```


### Query and joint data
We can joint PhytoestrogensUrine, PhthalatesUrine, and DietaryInterviewTotalNutrientIntakesFirstDay along with BodyMeasures, and BloodPressure. we got some data for those participants from the year 2003 to 2008 and attended all of those survey. 


```{r query2,warning=FALSE,message=FALSE}
df1 <- query_joint_data(config = "./phenotypes3.json")
nrow(df1)
unique(df1$years)
```





### Check the data availability over the years
We have much more data if we just query and join PhytoestrogensUrine, PhthalatesUrine,BodyMeasures, and BloodPressure or join DietaryInterviewTotalNutrientIntakesFirstDay with BodyMeasures, and BloodPressure.

```{r query3,warning=FALSE,message=FALSE}
df2 <- query_joint_data(config = "./phenotypes4.json")
nrow(df2)
unique(df2$years)

df3 <- query_joint_data(config = "./phenotypes5.json")
nrow(df3)
unique(df3$years)

```



### Check Blood Pressure

In the following matrix, 0 means the tables have no data in that year, and 1 means they have data in corresponding years.
As we can see from the below, the number of missing diastolic and systolic values is the same for both the first and second reads.
Previously, we removed the diastolic values with 0 zeros. However, NHANES shows its value range is 0 to 120 mmHg, and I also found two paper that says it could 0. Therefore, I am not 0 diastolic means dead, and we should remove them. 

- "right ventricular pressure falls to its diastolic level of about [zero...."](https://www.sciencedirect.com/topics/engineering/diastolic-pressure), it says "ventricular pressure", so, I am not sure it the same thing we have in NHANES
- [“Zero” diastolic blood pressure](http://dx.doi.org/10.4103/0973-0311.183572)

```{r blood,warning=FALSE,message=FALSE}
nhanes_db <- DBI::dbConnect(RSQLite::SQLite(), "../nhanes.sqlite")
blood <- DBI::dbGetQuery(nhanes_db, "select  BPXDI1,BPXDI2, BPXSY1,BPXSY2 from BloodPressure")
DBI::dbDisconnect(nhanes_db)
sapply(blood, function(x)sum(is.na(x)))
sapply(blood, function(x)length(which(x==0)))
sapply(blood[blood$BPXDI1!=0,], function(x)length(which(x==0)))

```

