---
title: "NHANES testsuite"
output: rmarkdown::html_vignette
author: Deepayan Sarkar
date: "Updated on : `r date()`"
vignette: >
  %\VignetteIndexEntry{testsuite}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 1. Create Test Suites

We can add more test suites to test the bug-prone functions to ensure the functions work as expected.

First, we need to load `testthat` package as shown in the above with use  `use_test()` to create test suites under the `tests` folder in the package. To learn more details about test suites, you can read [R Packages (2e), chapter 15](https://r-pkgs.org/testing-design.html).
```{r setup}
library(phonto)
library(devtools)
library(testthat)
```

For example, the following command would create a test suite `tests/testhat/test-nhanesTables.R` with a simple template.
```{r,eval = FALSE}
# use_testthat() create tests/testhat...
use_test("nhanesTables")
```

## 2. Writing Test Suites

To test the functions, we often use `expec_*` functions (eg `expect_equal(), expect_gt(),expect_match()...`) and `test_that()` to verify the outputs. 

For example, we can check the column names and dimensions of `phonto::nhanesTables()` outputs are the same as `nhanesA::nhanesTables().`

```{r,error=TRUE}
ph = nhanesTables('EXAM', 2008)
test_that("nhanesTables returns a dataframe", {
  expect_type(dim(ph), "integer")
})
```


## 3. Analysis Findings and Fix Bugs (if needed)

 To verify if the tables can still be downloaded from the CDC website, we can try to download them.  
 
 Note: the download information may not show up if you run the code Rmarkdown or vignette page.
 
```{r,error=TRUE}
for(tb in diff_tables){
  download.file(paste0("https://wwwn.cdc.gov/Nchs/Nhanes/2007-2008/",tb,".XPT"),paste0(tb,".XPT"))
  file.remove(paste0(tb,".XPT")) # you can comments out this line to check the files
}

```
As we can see from the above results, all those tables are still downloadable from the CDC website. Therefore, we may consider `nhanesA` has some bugs that missed some tables.

