---
title: "Building Variables defs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Organizing Phenotypes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
# output: html_document
date: "2023-04-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get the data....
One of the important aspects of the PHESANT package is that they provided some characterization of all the phenotypes, thereby making analysis easier.

We aim to provide a similar capability for NHANES and this document provides both a description of what we did and a template for others to enhance, or correct our classifications.

We first load `phonto`, and extract all questions together with the variable name, the SasLabel etc.  In the first part of this vignette we then label the different variables that are probably not useful as phenotypes.  Examples are things like survey weights, comment fields, etc.

```{r cars}
  library(phonto)
 t1 = paste0("SELECT DISTINCT Questionnaire, Variable, Description, SaSLabel FROM  QuestionnaireVariables")
xx = phonto:::nhanesQuery(t1)
#make a vector to store our variable labels
outPut = rep(NA, nrow(xx))
apply(xx,2, function(x) sum(is.na(x)))
#Questionnaire      Variable   Description      SaSLabel 
#            0             0          4604          4437 
missLab = is.na(xx$SaSLabel) |is.na(xx$Description)
usedIndex = which(missLab)
outPut[missLab] = "Missing"
```

## Any survey weights


And we can see already that some of the variables that start with WT have either a missing Description field or a missing SasLabel field.
```{r weights}
 isWT = grep("^WT", xx$Variable)
 outPut[isWT] = "Survey Weight"
 table(outPut, useNA="always")
 
 ##variance estimates also survey weights
 isWT = grep("^SDMV", xx$Variable)
 outPut[isWT] = "Survey Weight"
 
 ##now for Interview questions - eg what language was it done in etc
 ##unlikely to be relevant, but maybe
 
 g1 = grep("*[Ii]nterview\\??$", xx$SaSLabel)
 outPut[g1] = "Interview"
 
 ##comments - most variables end with an LC but not all...so we also
 ## look at the description - the SasLabel doesn't work as they abbreviate
 ##comments in weird ways
 g1 = grep("LC$", xx$Variable)
 g2 = grep("*[Cc]omment [Cc]ode$", xx$Description)
 outPut[union(g1,g2)] = "Comment"
 
```

##Build a searchable corpus

```{r eval=FALSE}
library("corpustools")
nhanes_df = xx
nhanes_df$Unique = paste0(nhanes_df$Questionnaire,"_", nhanes_df$Variable)
nhanes_tc = create_tcorpus(nhanes_df, doc_column = 'Unique', text_columns = 'Description')
nhanes_tc$preprocess(use_stemming = TRUE, remove_stopwords=TRUE)

h1 = search_features(nhanes_tc, query = c(`"blood pressure"`))

h2 = search_features(nhanes_tc, query = "hypertension")

  save=FALSE
  if( save ) {
   save(efo_tc, file= paste0(path, "/nhanes_tc.rda"), compress="xz")
   save(efo_df, file=paste0(path, "/nhanes_df.rda"), compress="xz")
  }


```
 And more from DEMO_I - not clear but they don't seem to be useful phenotypes

 SDDSRVYR - Data release cycle...
 RIDSTATR - Interview/Examination status
 SIAPROXY - Proxy used in SP Interview?
 SIAINTRP - Interpreter used in SP Interview?

